---
title: "Final Project Outline"
author: "Coco Kusiak, Joshua Freeman, and Luke Toomey"
date: "11/14/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(plyr)
require(mosaic)
```


```{r loading_the_data, echo = FALSE}
mhfp <- read.csv("bstat743fp.csv")
mhfp <- plyr :: rename(mhfp, c("SLEPTIM1"="sleep", "SEX" = "sex", "EMPLOY1"="employ", "SXORIENT"="sexorient",  
                       "TRNSGNDR"="trans", "QLACTLM2"="actlimit", "MSCODE"="metro", "X_TOTINDA"="exer30", 
                       "X_RACE_G1"="race", "X_RFBMI5"="bmi",  "X_INCOMG"="income", "X_SMOKER3"="smoker",
                       "X_RFBING5"="binge", "X_LLCPWT"="weight"))
mhfp <- mutate(mhfp, male = ifelse(sex == 1, yes = 1, no = ifelse(sex == 2, yes = 0, no = NA)),
               sleep = ifelse(sleep == 99 | sleep == 77, yes = NA, no = sleep),
               sexorient = ifelse(sexorient == 1, yes = "straight", 
                                  no = ifelse(sexorient == 2, yes = "gay", 
                                              no = ifelse(sexorient == 3, "bisexual", 
                                                          no = ifelse(sexorient == 4, yes = "other", no = NA)))),
               exer30 = ifelse(exer30 == 1, yes = 1, 
                               no = ifelse(exer30 == 2, yes = 0, no =NA)),
               trans = ifelse(trans %in% c(1, 2, 3), yes = 1, 
                              no = ifelse(trans == 4, yes = 0, no = NA)),
               smoker = ifelse(smoker %in% c(1, 2), yes = 1, 
                               no = ifelse(smoker == 3, yes = 2, 
                                           no = ifelse(smoker == 4, yes = 0, no =NA))),
               income = ifelse(income == 9, yes = NA, no = income),
               actlimit = ifelse(actlimit %in% c(7, 9 ), yes = NA, no = actlimit),
               employ = ifelse(employ == 9, yes = NA, no = employ),
               bmi = ifelse(bmi == 1, yes = 0, no = ifelse(bmi == 2, yes = 1, no = NA)),
               binge = ifelse(binge == 1, yes = 0, no = ifelse(binge == 2, yes = 1, no =NA))
)
```  



#The Data

-subset of BRFSS data


#The Variables

##The Outcome  

QLMENTL2: felt depression symptoms in the last 30 days


##The Covariates  
  -binge: binary for binge drinking in last 30 days
  -exer30: binary for exercising in the last 30 days, 794 missing
  -metro: 4 levels of metropolitan status, 240,000 missing
  -race: 5 level race variable
  -income: 5 levels of income, 81,000 missing
  -employ: 9 levels employment status, 73 missing
  -sex: gender
  -sexorient: sexual orientation, 280,000 missing
  -trans: 6 level, 281,000 missing
  -smoker: 4 level smoking status, 19,000 missing
  -sleep: continuous number of hours of sleep within 24 hour period, 2 missing
  -bmi: Overweight or obese calculated variable
  -actlimit: Activity Limitation Due to Health Problems

 

 
#Individual Project: Josh  
 
 
#Individual Project: Coco
I plan on comparing LASSO, ridge regression, and step-wise regression methods for model selection. For LASSO and ridge regression, I plan on using the `glmnet` package. For step-wise selection, I'll use the `stepAIC()` function in the `MASS` package. I plan on comparing the performance between these three models using 10-fold cross validation. 



#Individual Project: Luke
